---
title: "2. Exploratory Factor Analysis: Personality Traits"
author: "Sirid Wihlborg"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, psych)
```

# Table of Content
1. Data Cleaning
2. Create correlation matrix
3. perform PCA
4. Merge with ID
5. Display and Export new dataframe to directory

# 1. Data Cleaning
```{r}
# Loading the full dataset
data_big5_raw <- read_csv("../data/StudentLife/dataset/survey/BigFive.csv")


# selecting only the "types=pre"
data_big5 <- data_big5_raw %>% filter(type == "pre")

# disselecting the two first demographic rows
data_big5 <- data_big5 %>% select(-1, -2)

  
# find unique values in columns 
unique(data_big5$`I see myself as someone who...   - 1. Is talkative`)

# replacing character string with numeric value
data_big5 <- data_big5 %>% mutate_all(funs(if_else(. == "Disagree Strongly", "1", .)))
data_big5 <- data_big5 %>% mutate_all(funs(if_else(. == "Disagree a little", "2", .)))
data_big5 <- data_big5 %>% mutate_all(funs(if_else(. == "Neither agree nor disagree", "3", .)))
data_big5 <- data_big5 %>% mutate_all(funs(if_else(. == "Agree a little", "4", .)))
data_big5 <- data_big5 %>% mutate_all(funs(if_else(. == "Agree strongly", "5", .)))

# convert all values from character to numeric
data_big5 <- data_big5 %>% mutate_all(funs(as.numeric(.)))

# replace all missing values with 3  (this shoud be discussed)
sum(is.na(data_big5)) # first count amount of missing values 

data_big5 <- data_big5 %>% mutate_all(funs(if_else(is.na(.), 3, .)))

sum(is.na(data_big5)) # make sure there are none left
```


# 2. Create correlation matrix
```{r}
# create a heatmap
heatmap(cor(data_big5))
```
# 3. perform PCA
```{r}
fa.parallel(data_big5, fa="fa") # scree plot # 9 factors
fa_big5 <- fa(data_big5, 5, rotate="varimax")
print(fa_big5$loadings, cutoff=0.4)
```

MR2: 
Does a thorough job                            0.690
Is a reliable worker                           0.787
Tends to be disorganized                      -0.615
Tends to be lazy                              -0.694
Perseveres until the task is finished          0.654
Does things efficiently                        0.535
Makes plans and follows through with them      0.707
Is easily distracted                          -0.570

MR1: 
Tends to find fault with others                 0.658
Is relaxed, handles stress well.               -0.588
Worries a lot                                  0.680
Is emotionally stable, not easily upset       -0.508
Can be moody                                   0.700


MR5: 
Is talkative                                   -0.614
Is reserved                                     0.797
Tends to be quiet                              0.799
Has an assertive personality                  -0.532
Is sometimes shy, inhibited                    0.679
Is outgoing, sociable                         -0.688


MR3: 
Is original, comes up with new ideas            0.806
Is curious about many different things         0.580
Is ingenious, a deep thinker                   0.608
Has an active imagination                      0.607

MR4:
Is helpful and unselfish with others            0.566
Has a forgiving nature                         0.587
Is generally trusting                          0.530
Values artistic, aesthetic experiences         0.619
Is considerate and kind to almost everyone     0.780



The Big 5 Trades are: Openness, Conscientiousness, Extraversion, Agreeableness, and Neuroticism

MR1 = Neuroticism
MR2 = Conscientiousness
MR3 = Openness
MR4 = Agreeableness
MR5 = Extraversion (inverted)


# 4. Merge with ID
```{r}
fa_big5$scores

df_big5 <- cbind(data_big5_raw$uid, fa_big5$scores)
df_big5 <- as.data.frame(df_big5)
df_big5[,-1] <- lapply(df_big5[,-1], as.numeric) # make all values numeric (except the ID)

df_big5 <- rename(df_big5, "ID" = "V1", "Neuroticism" = "MR1", "Conscientiousness" = "MR2", "Openness" = "MR3", "Agreeableness" = "MR4", "Extraversion" = "MR5")

# obs: as the Extraversion column is inverted, we'll just changed the signs of the values in that column
df_big5$Extraversion <- (df_big5$Extraversion)*-1
```

# 5. Display and Export new dataframe to directory
```{r}
df_big5

write.csv(df_big5, "../df/df_big5.csv")
```







