---
title: "1. Preprocessing"
author: "Sirid Wihlborg"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse)
```

# Table of Content

1. Description of data
2. Cleaning the academic performance data
3. Creating a single measure for perceived stress level

## 1. Description of data

All the data is contained within the folder: "dataset" with the path: ../data/StudentLife/dataset. This folder includes: EMA, calendar, dinning, sensing, survey, app_usage, call_log, education, sms.


## 2. Cleaning the academic performance data

```{r}
# Load the data
data_grades_raw <- read_csv("../data/StudentLife/dataset/education/grades.csv")

# selecting the two first columns
df_grades <- data_grades_raw %>% select(1, 2)

# renaming columns
df_grades <- rename(df_grades, "ID" = "uid", "GPA" = `gpa all`)

# Display and Export new data frame to directory
df_grades

write.csv(df_grades, "../df/df_grades.csv")
```

## 3. Creating a single measure for perceived stress level

```{r}
# Loading the full dataset
data_stress_raw <- read_csv("../data/StudentLife/dataset/survey/PerceivedStressScale.csv")

# selecting only the "types=pre"
data_stress <- data_stress_raw %>% filter(type == "pre")

# find unique values in columns 
unique(data_stress$`1. In the last month, how often have you been upset because of something that happened unexpectedly?`)

# replace all missing values with "Sometime" (this should be discussed)
data_stress <- data_stress %>% mutate_all(funs(if_else(is.na(.), "Sometime", .)))
sum(is.na(data_stress)) # make sure there are no missing values left
```

Obs: before continuing, it's important to check the "direction" of the questions, ie., is the question phrased as such that an "agree" indicates a high or low stress level?
Based on visually inspecting the columns, it is found that for question 1,2,3,6,9,10 an "agree" would indicate high stress level, where for question 4,5,7,8 it would indicate low stress level. 

Theefor the values has been converted into numeric values in the following way: 

```{r}
# converting to numeriv values for the questions in one direction
data_stress <- data_stress %>% mutate_at(vars(3,4,5,8,11,12), funs(if_else(. == "Never", "1", .))) # obs: remember that the first two columns are demographic info
data_stress <- data_stress %>% mutate_at(vars(3,4,5,8,11,12), funs(if_else(. == "Almost never", "2", .)))
data_stress <- data_stress %>% mutate_at(vars(3,4,5,8,11,12), funs(if_else(. == "Sometime", "3", .)))
data_stress <- data_stress %>% mutate_at(vars(3,4,5,8,11,12), funs(if_else(. == "Fairly often", "4", .)))
data_stress <- data_stress %>% mutate_at(vars(3,4,5,8,11,12), funs(if_else(. == "Very often", "5", .)))

# converting to numeriv values for the questions in the other direction
data_stress <- data_stress %>% mutate_at(vars(6,7,9,10), funs(if_else(. == "Never", "1", .)))
data_stress <- data_stress %>% mutate_at(vars(6,7,9,10), funs(if_else(. == "Almost never", "2", .)))
data_stress <- data_stress %>% mutate_at(vars(6,7,9,10), funs(if_else(. == "Sometime", "3", .)))
data_stress <- data_stress %>% mutate_at(vars(6,7,9,10), funs(if_else(. == "Fairly often", "4", .)))
data_stress <- data_stress %>% mutate_at(vars(6,7,9,10), funs(if_else(. == "Very often", "5", .)))

# convert all values from questions from character to numeric
data_stress <- data_stress %>% 
  mutate_at(vars(c(3:12)), funs(as.numeric(.)))
```

```{r}
# create a column with the mean score
data_stress <- data_stress %>%
  mutate(Perceived_Stress = rowMeans(data_stress[, 3:12]))

# renaming columns
data_stress <- rename(data_stress, "ID" = "uid")

# selecting columns to use
df_stress <- data_stress %>% select("ID", "Perceived_Stress")

# exporting data frame
write.csv(df_stress, "../df/df_stress.csv")
```




