---
title: "3. Correlation Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Table of Content 

1. Load data and packages
2. Visualise correlations  
    2.1 Big5 and Grades  
    2.2 Stress and Grades  
3. Combined Plots  
    3.1 Restructure data  
    3.2 Big5 and GPA  
4. Linear Regression

## 1. Load data and packages
```{r}
pacman::p_load(tidyverse, reshape2, RColorBrewer)

df_big5 <- read_csv("../df/df_big5.csv")
df_stress <- read_csv("../df/df_stress.csv")
df_grades <- read_csv("../df/df_grades.csv")

df <- merge(df_big5, df_stress, by = "ID")
df <- merge(df, df_grades, by = "ID")

# removing unnecessary columns
df <- df %>% select(-2,-8,-10)
```

## 2. Visualise correlations

### 2.1 Big5 and Grades

```{r}
# Create a vector of column names to loop through
columns <- c("Extraversion", "Agreeableness", "Neuroticism", "Openness", "Conscientiousness")

# Loop through the columns
for (col in columns) {
  
  # Plot the data
  df %>% 
    ggplot(aes_string(col, "GPA")) + 
    geom_point() +
    geom_smooth(method = "lm", se = FALSE, color = "blue", size = 1.5) + 
    ggtitle(paste(col, "and GPA")) +
    labs(x = col, y = "GPA") +
    theme_minimal() +
    theme(plot.background = element_rect(fill = "white"),
          text = element_text(size = 14),
          axis.ticks = element_line(color = "grey"))
  
  # save the plots to directory
  ggsave(paste0("plots/GPA_", col, ".png"), width = 8, height = 6)
}
```

### 2.2 Stress and Grades

```{r}
# Plot the data
df %>% 
  ggplot(aes(Perceived_Stress, GPA)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "orange", size = 1.5) + 
  ggtitle("Perceived Stress and GPA") +
  labs(x = "Perceived Stress", y = "GPA") +
  theme_minimal() +
  theme(plot.background = element_rect(fill = "white"),
        text = element_text(size = 14),
        axis.ticks = element_line(color = "grey"))

# save the plots to directory
ggsave(paste0("plots/GPA_Stress.png"), width = 8, height = 6)

```

## 3. Combined Plots

### 3.1 Restructure data
```{r}
# Reshape the data frame into a long format
df_long <- melt(df, id.vars = c("ID", "GPA", "Perceived_Stress"), measure.vars = c("Extraversion", "Agreeableness", "Neuroticism", "Openness", "Conscientiousness"))
```



### 3.2 Big5 and GPA
```{r}
df_long %>% 
  ggplot(aes(value, GPA)) +
    geom_smooth(aes(group = variable, color=variable), method = "lm", se = FALSE) +
    scale_color_brewer(palette = "Dark2") +
    theme_minimal() + 
    ylim(2.8,3.9) + 
    ggtitle("Correlation between Academic Performance and Personality Trades") +
    labs(
      subtitle = "GPA Score and Personality Trades extracted fromt the Big 5 questionnaire",
      x = "Trade Score", 
      y = "GPA", 
      color = "Big 5 Trade:") +
    theme(plot.background = element_rect(fill = "white"),
          text = element_text(size = 12),
          axis.ticks = element_line(color = "grey"))

# save the plot
ggsave("plots/Combined_GPA_Big5.png", width = 8, height = 6)
```


## 4. Linear Regression

Modelling GPA as dependent on Personality traits & Perceived stress levels
```{r}
# Create a vector of column names to loop through
columns <- c("Extraversion", "Agreeableness", "Neuroticism", "Openness", "Conscientiousness", "Perceived_Stress")

# fitting the linear model as well as checking assumptions
for (col in columns) {
  # Create a formula using the paste() and as.formula() functions
  formula <- as.formula(paste("GPA ~", col))
  fit <- lm(formula, data=df)
  print(summary(fit))
  
  # checking the assumption of homoscedasticity
  plot(fitted(fit), residuals(fit), xlab = "Fitted values", ylab = "Residuals")
  abline(h = 0, col = "red")
  
  # checking the assumption of normality
  qqnorm(residuals(fit), main = "Q-Q Plot of Residuals")
  qqline(residuals(fit))
}


```





